A user-extensible and adaptable parser architecture

### 2008-05-27 papers/sgai-2008/paper.tex
Added a chunk of logparser.pdf's coverage section.

>>>>

The discussion of the parser's coverage of Postfix log files is separated
into two parts: log lines covered and mails covered.  The first is
important because the parser should handle all (relevant) log lines it's
given; the second is equally important because the parser must properly
deal with every mail if it is to be useful.  Improving the former is
less intrusive, as it just requires new rules to be written; improving the
latter is much more intrusive as it requires changes to the parser
algorithm, and it can also be much harder to notice a deficiency.

Warnings are issued for any log lines which are not parsed; no warnings are
issued for unparsed log lines while testing with the \numberOFlogFILES{}
test log files, so it can be safely concluded that there are zero false
negatives.  False positives are harder to quantify: short of examining each
of the 60,721,709 log lines and determining which \regex{} parsed it, there
is no way to be sure that every line was parsed by the correct \regex{},
making it impossible to quantify the false positive rate; however a random
sample of 6039 log lines was parsed and the results checked manually to
ensure that the correct \regex{} parsed each log line.

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails which were not handled properly.  The parser
issues warnings when it detects any errors, some of which may alert the
user to a problem.  There should be few or no warnings when parsing, and
when finished parsing the state table should only contain entries for mails
which had yet to be delivered when the log files ended, or were accepted
before the log files began.

Verifying by inspection that the parser correctly deals with all 60,721,709
lines in the test log files is infeasible, but verifying a subset of those
log files is a tractable, if extremely time consuming, task.  A sample of
log lines was parsed with all debugging options enabled, resulting in
167,448 lines of output.  All 167,448 lines were examined in conjunction
with the input log and a dump of the resulting database, verifying that for
each of the input lines the parser used the correct rule and executed the
correct action, which in turn produced the correct result and inserted the
correct data in the database.  The log segment produced 4 warnings, 10
mails remaining in the state tables, and 1625 connections correctly entered
in the database.

Given the evidence detailed above, the author is confident that the false
positive rate when reconstructing a mail is exceedingly low, if not
approaching zero.

Parser coverage is divided into two topics in this section: log lines
covered and mails covered.\footnote{There is substantially more coverage
discussion in the author's thesis.}  The former is initially more
important, as the parser must successfully parse every line if it is to be
complete, but subsequently the latter takes precedence because reproducing
the path a mail takes through Postfix is the aim of the parser.  Increasing
the percentage of log lines parsed is relatively simple and non-intrusive:
adding new rules or modifying existing rules is simplified by the
separation of rules, actions and framework.  Improving the logical coverage
is harder, as the actions taken by Postfix must be reconstructed by the
author, and the new sequence of actions integrated into the existing model
without breaking the existing parsing.  Detecting a deficiency in the
parsing algorithm is also significantly harder than detecting unparsed log
lines, as the parser will warn about any unparsed line, whereas discovering
a flaw in the parser requires understanding of the warnings produced and
the mails remaining in the state table.  Rectifying a flaw in the parser
requires an understanding of both the parser and Postfix's log files, and
investigative work to determine the cause of the deficiency, followed by
further examination of the log files in developing a solution.
